---
title: "Stylistic Dispersion — Visualizations"
output:
  html_notebook:
    toc: true
    toc_float: true
---

Renders every visualization produced by `dispersion_regression.R` (the
primary analysis pipeline), plus a few additional views that summarize
findings which only existed as tables before now (the rating-level
heatmap in particular — see section 5).

Replaces `convergence_viz.R`, which was built around `convergence_fanfic.R`'s
outputs; that script and its brms cache files were deleted once the primary
analysis moved to the OLS + shrinkage approach here.

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.width = 10)
```

```{r load-pipeline}
# runs the full pipeline: data prep, all OLS fits, all three stage-2 pooling
# models. The three brms models load from outputs/brms_cache/ if unchanged
# (a few seconds); only refits if the cache is stale or missing.
source("dispersion_regression.R")
```

## 1. Headline: dispersion trend, pooled

The main result. Fandom x factor, colored by the pooled (shrunk) time-slope;
starred where the 95% credible interval excludes zero. `random` is excluded
from the pooling (shown separately in the console output as an unpooled
control) — see the script for why.

```{r heatmap-main, fig.height=4}
plot_dispersion_heatmap(shrunk)
```

## 2. Raw vs pooled: how much did shrinkage change each estimate?

Open grey circle = raw stage-1 OLS slope; filled point + bar = pooled
estimate with 95% CrI. Segment length shows how far partial pooling pulled
each raw number — long segments are exactly the cells you should be most
skeptical of if you only had the raw estimate.

```{r forest-main, fig.height=11}
plot_shrinkage_forest(shrunk)
```

## 3. Per-factor detail: binned dispersion over time

Binned mean |deviation from fandom median| against time, per fandom, with a
loess smooth. Point size = bin n, so precision is visible directly. Useful
for spotting non-linear patterns a linear `time` slope would average away
(e.g. RPF's F1 trend is U-shaped, not monotonic — worth remembering when
reading its "no credible change" result in the main table).

```{r trend-per-factor, fig.height=6, results='asis'}
for (fk in FACTORS) {
  cat("### ", fk, "\n")
  print(plot_dispersion_trend(d, fk))
  cat("\n\n")
}
```

## 4. Rating x time interaction: does rating change the RATE of change?

Fandom x factor, faceted by rating, colored by the pooled `rating:time`
interaction (shift relative to General Audiences' own trend). This is the
one that revealed RPF and PJ's divergence was concentrated in their
Explicit/Mature-rated content rather than uniform across ratings.

```{r heatmap-rating-time, fig.height=5, fig.width=12}
plot_rating_time_heatmap(shrunk_rt)
```

### Per-factor detail, raw estimates by fandom

Unpooled `rating:time` estimates (95% CI), one panel per factor, color =
fandom. Useful for seeing which specific fandom/rating pairs are driving
the pooled result above.

```{r rating-time-per-factor, fig.height=6, results='asis'}
for (fk in FACTORS) {
  cat("#### ", fk, "\n")
  print(plot_rating_time_by_fandom(rating_time_effects, fk))
  cat("\n\n")
}
```

## 5. Rating x average dispersion level

This is new — the additive `rating` effect (does dispersion sit higher or
lower overall for this rating, not its trend over time) was computed early
in the pipeline but never visualized until now. It turned out to be far more
pervasive than the interaction above (121 of 200 cells credible) and shows a
clean split: F3/F4/F7/F8 (Negative Affect, length, Conversationality,
Positive Affect) sit **lower** in rated content across nearly every fandom;
F1/F9/F10 (Syntactic Complexity, Narrative Drift, Structural Variability)
sit **higher**.

```{r heatmap-rating-level, fig.height=5, fig.width=12}
plot_rating_level_heatmap(shrunk_rl)
```

## 6. Other metadata: logwords / growth_resid / is_new_author

Not pooled (nuisance covariates, not a primary question) — just a quick
visibility pass so these aren't left completely unexamined. `logwords` is a
strong, expected control (longer texts deviate less from the median);
`growth_resid` contributes little on its own; `is_new_author` shows a
modest, plausible "newcomers are somewhat more idiosyncratic" pattern.

```{r other-metadata-plot, fig.height=5}
other_metadata_summary %>%
  mutate(term = factor(term, levels = c("logwords", "growth_resid", "is_new_authorTRUE"))) %>%
  ggplot(aes(fandom, mean_estimate, fill = n_credible)) +
  geom_col() +
  geom_hline(yintercept = 0, linetype = 2, color = "grey40") +
  facet_wrap(~ term, scales = "free_y") +
  scale_fill_gradient(low = "grey80", high = "#d7191c", name = "n factors\ncredible") +
  labs(title = "Other metadata effects on dispersion, mean estimate per fandom",
       x = NULL, y = "mean coefficient across 10 factors") +
  theme_bw(base_size = 11) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

## 7. Non-linearity robustness check

Motivated by looking at section 3's trend plots — MAG and PJ's F8 both show
what looks like a rise-then-decline shape, not a straight line. For every
credible cell in `shrunk`, refit with a flexible (natural spline) time term
and check whether that curvature is statistically real, and whether the
early-vs-late direction still agrees with the linear model's label.

**Result: none of the 14 credible cells show significant non-linearity**
(smallest p = .094, MAG-F8), and 13 of 14 keep the same direction under the
flexible fit. The visual "hump" in MAG/PJ-F8 below is real in the point
estimate, but the confidence ribbon is wide enough that a straight line
stays plausible inside it — the linear results hold up.

```{r nonlinearity-table}
nonlinearity_check %>% arrange(nonlinear_p)
```

```{r spline-comparison, fig.height=6}
print(plot_spline_comparison(d, "MAG", "F8"))
print(plot_spline_comparison(d, "PJ", "F8"))
```

## 8. Collinearity checks (diagnostic, not a finding)

`growth_resid` is orthogonalized to `time` by construction — confirms it
stayed that way for the OLS predictor set (section 6 of the script) and for
the rating interaction's predictor set (section 7's `rating_time_collinearity`).

```{r vif-table}
vif_by_fandom
```

```{r rating-collinearity-table}
rating_time_collinearity %>% arrange(desc(incremental_r2_from_rating))
```
